Fast Algorithms for Intersection of Non-matching Grids Using \plucker Coordinates.

### 2016-12-20 paper_intersections.tex
Small typos and fixes, remove initials at references.

According to our knowledge, there are lots of works using non-matching grids, yet only few of them discuss algorithms how to compute their intersections. 
Gander and Japhet \cite{gander_algorithm_2013} present the PANG algorithm for 2d-2d and 3d-3d intersections that can be used e.g. for mesh overlapping methods.
They use the advancing front technique to get candidate pairs in linear time. The algorithm is part of the library DUNE \cite{bastian_towards_2005}.
Massing, Larson, and Logg \cite{massing_efficient_2013} present an algorithm for 2d-3d intersections as part of their implementation of the Nitche method 
which is part of the project Dolfin \cite{Dolphin}. They use axes aligned bounding boxes of elements (AABB) and bounding interval hierarchy (BIH) 
to get intersection candidate pairs of elements, the GTS library \cite{gtslib} is used for 2d-3d intersections. 
Finally, there is the work of A. Elsheikh and M. Elsheikh \cite{elsheikh_reliable_2012} presenting an algorithm for 2d-2d mesh union operation which includes 
calculation and imprinting of the intersection curves. They exploit the binary space partitioning for searching of the initial intersection and 
the advancing front method for the intersection curve tracking.

In this paper we present a new approach to the mesh intersection problem based on the \plucker coordinates,
further developing the algorithm of Platis and Theoharis \cite{platis_fast_2003} for ray-tetrahedron intersections. 
Presented intersection algorithms for pairs of simplicial elements of different dimensions are based on \plucker coordinates.
These algorithms are combined with the advancing front method which allows us to reuse \plucker coordinates and their products between neighboring
elements and reduce the number of arithmetic operations.

@@

\subsection{\plucker Coordinates}
\plucker coordinates represent a line in 3d space. The definition properties and more general context from computational 
geometry can be found e.g. in \cite{dorst_geometric_2007} or \cite{joswig_polyhedral_2013}.
Considering a line $p$, given by a point $\vc A_p$ and its directional vector $\vc{u}_p$, 
the \plucker coordinates of $p$ are defined as
\[ \pi_p = (\vc{u}_p, \vc{v}_p),\; \vc{v}_p = \vc{u}_p\times \vc A_p. \]
This representation is independent on choice of $A_p$ since $\vc{u}_p\times (\vc A_p + t \vc u) = \vc{u}_p\times \vc A_p$.
Further, having two lines $p$ and $q$ with \plucker coordinates $\pi_p$ and $\pi_q$, we denote a permuted inner product
\[\pi_p \odot \pi_q = \vc{u}_p\cdot \vc{v}_q + \vc{u}_q \cdot \vc{v}_p. \]

The sign of the permuted inner product is non-zero if $p$ and $q$ are skew lines and is positive if $q$ is oriented counterclockwise and 
negative if $q$ is oriented clockwise looking in the $p$ direction. This can be used to determine relative position 
of the line $p$ and the triangle. This is demonstrated in \fig{fig:plucker_products}.
The permuted inner products of triangle sides with the line $p$ have common sign, cases (a) and (c), if and only if the 
line intersect the triangle inside. If any $\pi_p\odot \pi_{s_i}$ is zero, as in the case (b), 
it means that the lines $p$ and $s_i$ are coplanar.

\begin{figure}[!htb]
    \centering
    \subfloat[$\pi_p\odot \pi_{s_i} < 0 \;\forall i$]{
        \includegraphics[width=0.25\textwidth]{\figpath plucker_product_a.pdf}
    }
    \hspace{2ex}
    \subfloat[$\pi_p\odot \pi_{s_i} = 0 \, \exists i$, \newline special case]{
        \includegraphics[width=0.25\textwidth]{\figpath plucker_product_b.pdf}
    }
    \hspace{2ex}
    \subfloat[$\pi_p\odot \pi_{s_i} > 0 \;\forall i$]{
        \includegraphics[width=0.25\textwidth]{\figpath plucker_product_c.pdf}
    }
    \caption{Different relative positions of the line $p$ and a triangle with sides $s_i$, $i=0,1,2$. 
            Dashed parts are behind the triangle. Signs of the permuted inner products depend on orientation of lines,
            the line $p$ is coplanar with some side in the case (b).
    }
    \label{fig:plucker_products}
\end{figure}

@@

\subsection{Intersection Line-Triangle (1d-2d)}
\label{sec:1d-2d}
Let us consider a line $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,
\end{equation}
on which a line segment $S_1$ is defined by $t\in [0,1]$ and a triangle $S_2$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$, see \fig{fig:triangle_notation}

>>>>

According to our knowledge, there are lots of works using non-matching grids, yet only few of them discuss algorithms how to compute their intersections. 
Gander and Japhet \cite{gander_algorithm_2013} present the PANG algorithm for 2d-2d and 3d-3d intersections that can be used e.g. for mesh overlapping methods.
They use the advancing front technique to get candidate pairs in linear time. The algorithm is part of the library DUNE \cite{bastian_towards_2005}.
Massing, Larson, and Logg \cite{massing_efficient_2013} present an algorithm for 2d-3d intersections as part of their implementation of the Nitche method 
which is part of the project Dolfin \cite{Dolphin}. They use axes aligned bounding boxes of elements (AABB) and bounding interval hierarchy (BIH) 
to get intersection candidate pairs of elements, the GTS library \cite{gtslib} is used for 2d-3d intersections. 
Finally, there is the work of Elsheikh and Elsheikh \cite{elsheikh_reliable_2012} presenting an algorithm for 2d-2d mesh union operation which includes 
calculation and imprinting of the intersection curves. They exploit the binary space partitioning for searching of the initial intersection and 
the advancing front method for the intersection curve tracking.

In this paper we present a new approach to the mesh intersection problem based on the \plucker coordinates,
further developing the algorithm of Platis and Theoharis \cite{platis_fast_2003} for ray-tetrahedron intersections. 
Presented intersection algorithms for pairs of simplicial elements of different dimensions are based on \plucker coordinates.
These algorithms are combined with the advancing front method which allows us to reuse \plucker coordinates and their products among neighboring
elements and reduce the number of arithmetic operations.

@@

\subsection{\plucker Coordinates}
\plucker coordinates represent a line in 3d space. The definition properties and more general context from computational 
geometry can be found e.g. in \cite{dorst_geometric_2007} or \cite{joswig_polyhedral_2013}.
Considering a line $p$, given by a point $\vc A_p$ and its directional vector $\vc{u}_p$, 
the \plucker coordinates of $p$ are defined as
\[ \pi_p = (\vc{u}_p, \vc{v}_p),\; \vc{v}_p = \vc{u}_p\times \vc A_p. \]
This representation is independent of the choice of $A_p$ since $\vc{u}_p\times (\vc A_p + t \vc u) = \vc{u}_p\times \vc A_p$.
Further, having two lines $p$ and $q$ with \plucker coordinates $\pi_p$ and $\pi_q$, we denote a permuted inner product
\[\pi_p \odot \pi_q = \vc{u}_p\cdot \vc{v}_q + \vc{u}_q \cdot \vc{v}_p. \]

The sign of the permuted inner product is non-zero if $p$ and $q$ are skew lines and is positive if $q$ is oriented counterclockwise and 
negative if $q$ is oriented clockwise looking in the $p$ direction. This can be used to determine relative position 
of the line $p$ and the triangle. This is demonstrated in \fig{fig:plucker_products}.
The permuted inner products of triangle sides with the line $p$ have common sign, cases (a) and (c), if and only if the 
line intersects the triangle inside. If any $\pi_p\odot \pi_{s_i}$ is zero, as in the case (b), 
it means that the lines $p$ and $s_i$ are coplanar.

\begin{figure}[!htb]
    \centering
    \subfloat[$\pi_p\odot \pi_{s_i} < 0 \;\forall i$]{
        \includegraphics[width=0.25\textwidth]{\figpath plucker_product_a.pdf}
    }
    \hspace{2ex}
    \subfloat[$\pi_p\odot \pi_{s_i} = 0 \, \exists i$, \newline special case]{
        \includegraphics[width=0.25\textwidth]{\figpath plucker_product_b.pdf}
    }
    \hspace{2ex}
    \subfloat[$\pi_p\odot \pi_{s_i} > 0 \;\forall i$]{
        \includegraphics[width=0.25\textwidth]{\figpath plucker_product_c.pdf}
    }
    \caption{Different relative positions of the line $p$ and a triangle with sides $s_i$, $i=0,1,2$. 
            Dashed parts are behind the triangle. Signs of the permuted inner products depend on orientation of lines,
            the line $p$ is coplanar with a side in the case (b).
    }
    \label{fig:plucker_products}
\end{figure}

@@

\subsection{Intersection Line-Triangle (1d-2d)}
\label{sec:1d-2d}
Let us consider a line $p$ with parametric equation 
\begin{equation}
    \label{eq:line_parametric}
    \vc X = \vc A + t\vc u,
\end{equation}
on which a line segment $S_1$ is defined by $t\in [0,1]$ and a triangle $S_2$ given by vertices $(\vc V_0, \vc V_1, \vc V_2)$ 
with oriented sides $s_i=(\vc V_j, \vc V_k)$, $j=(i+1)\text{ mod }3$, $k=(i+2)\text{ mod }3$, see \fig{fig:triangle_notation}.
